Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
got rid of exploration terminology

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a neural network (recognition model). Each of these steps iteratively bootstraps off of the others.}  \label{feeding}\vspace{-0.4cm}
\end{wrapfigure}
  \end{comment}
\\\noindent \textbf{Exploration: Searching for programs.}  Our program search
is informed by both the DSL and the recognition model. When these
improve, we can solve more tasks.
\\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
from the programs found in the exploration phase, aiming
to maximally compress (or, raise the prior probability of) these
programs.  As we solve more tasks, we hone in on DSLs that
more closely match the domain.
\\\noindent \textbf{Compilation: Learning a neural recognition model.}  We
update the recognition model by training on two data sources: samples
from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
programs found by the search procedure during exploration. As the DSL
improves and as search finds more programs, the recognition model gets
more data to train on, and better data.

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Exploration) and the DSL (Compression):
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\subsection{Exploration: Searching for Programs}\label{explorationSection}

@@

Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
We note the following implementation details:
(1) We perform an exploration cycle before each compression and compilation  cycle.
(2) On the first iteration, we do \emph{not} train the recognition model on samples from
the generative model because a generative model has not yet been learned --
we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
%% (3) Because the frontiers can grow very large,
%% we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
%% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
(3) During both DSL induction and neural net training, we calculate Eq.~\ref{AIC} and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
(4) For added robustness, we enumerate programs from both the generative model $(\mathcal{D},\theta)$ and the recognition model.
\end{comment}

@@

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
We can not compare directly with these systems,
because they are engineered for one specific domain, and 
do not have publicly available code and datasets.
\\\noindent \textbf{EC}, an implementation of the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, publicly available at...
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first exploration step.

>>>>

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Exploration)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Exploration)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a neural network (recognition model). Each of these steps iteratively bootstraps off of the others.}  \label{feeding}\vspace{-0.4cm}
\end{wrapfigure}
  \end{comment}
\\\noindent \textbf{Search: Solving tasks.}  Our program search
is informed by both the DSL and the recognition model. When these
improve, we can solve more tasks.
\\\noindent\textbf{Compression: Improving the DSL.} We induce the DSL
from the programs found in the search phase, aiming
to maximally compress (or, raise the prior probability of) these
programs.  As we solve more tasks, we hone in on DSLs that
more closely match the domain.
\\\noindent \textbf{Compilation: Learning a neural recognition model.}  We
update the recognition model by training on two data sources: samples
from the DSL (as in the Helmholtz Machine's ``sleep'' phase), and
programs found by the search procedure during search. As the DSL
improves and as search finds more programs, the recognition model gets
more data to train on, and better data.

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Search) and the DSL (Compression):
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\subsection{Searching for Programs}\label{explorationSection}

@@

Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
We note the following implementation details:
(1) We perform an search cycle before each compression and compilation  cycle.
(2) On the first iteration, we do \emph{not} train the recognition model on samples from
the generative model because a generative model has not yet been learned --
we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
%% (3) Because the frontiers can grow very large,
%% we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
%% with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
(3) During both DSL induction and neural net training, we calculate Eq.~\ref{AIC} and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$ as measured by $\probability[x,p|\mathcal{D},\theta]$ -- we found that $K = 2$ sufficed.
(4) For added robustness, we enumerate programs from both the generative model $(\mathcal{D},\theta)$ and the recognition model.
\end{comment}

@@

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model.
\\\noindent \textbf{RF/DC}, which holds the generative model $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
We can not compare directly with these systems,
because they are engineered for one specific domain, and 
do not have publicly available code and datasets.
\\\noindent \textbf{EC}, an implementation of the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, publicly available at...
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning -- equivalently, our first search step.
